Accessibility Calculations

### 2023-07-06 Fifteen-minute-neighborhoods/Fifteen-minute-neighborhoods-title-page.Rmd
Updated analysis and files for paper.

bibliography: bibliography.bib
biblio-style: spphys
# bibstyle options spbasic(default), spphys, spmpsci
output: rticles::springer_article
---

>>>>

bibliography: 
  - '`r system(kpsewhich bibliography.bib)`'
biblio-style: spphys
# bibstyle options spbasic(default), spphys, spmpsci
output: rticles::springer_article
---

### 2023-07-06 data-raw/03-Network-Attributes-by-Walkshed.Rmd
Updated analysis and files for paper.

Load packages:
```{r warning=FALSE, message=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(igraph)
library(skimr) # Compact and Flexible Summaries of Data
library(sf) # Simple Features for R
library(sfnetworks)
library(tidyr)
```

@@

Motfis (from igraph help pages)
```{r}
g <- sample_pa(100)
motifs(g, 3)
count_motifs(g, 3)
sample_motifs(g, 3)

```{r}
pattern <- graph.full(3)
my.graph <- grg.game(100, 0.2) 
```

@@

>>>>

Load packages:
```{r warning=FALSE, message=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(igraph) # Network Analysis and Visualization
library(patchwork) # The Composer of Plots
library(skimr) # Compact and Flexible Summaries of Data
library(sf) # Simple Features for R
library(sfnetworks) # Tidy Geospatial Networks
library(tidyr) # Tidy Messy Data
```

@@

@@

Explore the motifs. First, extract a walkshed and convert it to sfnetworks class:
```{r}
sample_walkshed <- sample.int(nrow(walksheds_da), 1)

ggplot() +
  geom_sf(data = walksheds_da[sample_walkshed,]) +
  geom_sf(data = wsnet |>
            activate(edges) |>
            st_as_sf()
  ) +
  theme_void()

Note that there are 16 classes of directed isomorphic graphs of size 3. Here numbers 1 and 15 are obtained:
```{r}
# Size is the number of nodes, number is the isomorphic class, directed true
g1 <- graph_from_isomorphism_class(size = 3, number = 1, directed = TRUE)
g15 <- graph_from_isomorphism_class(size = 3, number = 15, directed = TRUE)
```

Plot isomorphism:
```{r}
ggraph(g1, layout = "linear", circular = TRUE) +
  geom_edge_fan2(arrow = arrow(length = unit(5, 'mm')),
                 end_cap = circle(5, 'mm'),
                 lineend = "round", 
                 width = 1) + 
  geom_node_point(size = 5) +
  coord_equal() +
  theme_void()

Obtain all directed isomorphisms of size 3:
```{r}
size_3_iso <- purrr::map(1:15,
                         ~graph_from_isomorphism_class(size = 3, 
                                                       number = .x, 
                                                       directed = TRUE))
```

```{r}
ggraph(size_3_iso[[10]], layout = "linear", circular = TRUE) +
  geom_edge_fan2(arrow = arrow(length = unit(5, 'mm')),
                 end_cap = circle(5, 'mm'),
                 lineend = "round", 
                 width = 1) + 
  geom_node_point(size = 5) +
  coord_equal() +
  theme_void()
```

Motifs of size 3 in wsnet:
```{r}
motifs(wsnet)
```

The most common by far is class 10.

```{r}
count_motifs(wsnet)
```

Initialize a data frame to save the motif counts:
```{r}
walksheds_da_motif_3_count <- data.frame(matrix(0, nrow = nrow(walksheds_da), ncol = 16))

walksheds_da_motif_3_count <- walksheds_da_motif_3_count |>
  rename_with(~ paste0("iso_class_", c(0:15)), starts_with("X"))

DA <- numeric(length = nrow(walksheds_da))
```

Compute the number of motifs of each class in the walksheds:
```{r warning=FALSE}
# Time algorithm
start.time <- Sys.time()

for(i in 1:nrow(walksheds_da)){
  wsnet <- hamilton_edges |>
    st_filter(
      walksheds_da[i,],
      .predicate = st_intersects)

  wsnet <- hamilton_edges |> 
    filter(edge_index %in% wsnet$edge_index)

  # Convert to sfnetwork
  wsnet <- as_sfnetwork(wsnet)
  walksheds_da_motif_3_count[i,] <- motifs(wsnet)
  DA[i] <- walksheds_da$GeoUID
}

walksheds_da_motif_3_count <- cbind(DA,
                                    walksheds_da_motif_3_count)

Time difference:
```{r}
time.taken
```

Joing urban fabric type:
```{r}
walksheds_da_motif_3_count <- walksheds_da_motif_3_count |>
  rename(GeoUID = DA) |>
  left_join(walksheds_da |>
              st_drop_geometry(),
            by = "GeoUID") |>
  select(GeoUID,
         Type,
         starts_with("iso_class"))
```

Save the data object:
```{r}
save(walksheds_da_motif_3_count, file = "walksheds_da_motif_3_count.rda", compress = "xz")
```

Summary of motif counts:
```{r}
summary(walksheds_da_motif_3_count)
```

The only motifs present are class 10 and class 15:
```{r}
iso_3_10_plot <- ggraph(size_3_iso[[10]], layout = "linear", circular = TRUE) +
  geom_edge_fan2(arrow = arrow(length = unit(5, 'mm')),
                 end_cap = circle(5, 'mm'),
                 lineend = "round", 
                 width = 1) + 
  geom_node_point(size = 5) +
  coord_equal() +
  theme_void()

iso_3_15_plot <- ggraph(size_3_iso[[15]], layout = "linear", circular = TRUE) +
  geom_edge_fan2(arrow = arrow(length = unit(5, 'mm')),
                 end_cap = circle(5, 'mm'),
                 lineend = "round", 
                 width = 1) + 
  geom_node_point(size = 5) +
  coord_equal() +
  theme_void()

iso_3_10_plot + iso_3_15_plot
```

Check walksheds with low number of motifs (normalized):
```{r}
d <- walksheds_da_net_vars |>
                  drop_na() |>
  mutate(normalized_motifs_3 = motifs_3/n_edges,
         normalized_motifs_4 = motifs_4/n_edges)

d |> filter(normalized_motifs_3 < 0.7)
```

Choose a walkshed with low number of motifs:
```{r}
ID <- "35250947" #"35250992"# "35250317"

wsnet <- hamilton_edges |>
  st_filter(
    walksheds_da |> filter(GeoUID == ID),
    .predicate = st_intersects)

# Convert to sfnetwork
wsnet <- as_sfnetwork(wsnet)
```

```{r}
ggplot()  +
  geom_sf(data = walksheds_da |>
            filter(GeoUID == ID),
          color = "blue",
          fill = "blue") +
  geom_sf(data = urban_hammer_da |> 
            filter(GeoUID == ID), 
          fill = "red") +
  geom_sf(data = wsnet |> 
            activate(edges) |> 
            st_as_sf())
```

Find the motifs of class 10 and 15 in wsnet:
```{r}
iso_class_10_list <- subgraph_isomorphisms(size_3_iso[[10]], 
                                           wsnet, 
                                           method = "vf2")

iso_class_15_list <- subgraph_isomorphisms(size_3_iso[[15]], 
                                           wsnet, 
                                           method = "vf2")
```

The elements in the list that results are triads of nodes. 
```{r}
subgraph(wsnet, vids = iso_class_10_list[[1]]) |>
  as_sfnetwork()
```

Extract the isomorphs of class 10 from the network:
```{r message=FALSE}
# This is the first one in the list
motifs_class_10 <- subgraph(wsnet, vids = iso_class_10_list[[1]]) |>
  as_sfnetwork()

# Join every other isomorph in the list
for(i in 2:length(iso_class_10_list)){
  motifs_class_10 <- st_network_join(motifs_class_10,
                   subgraph(wsnet, vids = iso_class_10_list[[i]]) |>
                     as_sfnetwork())
}
```

Extract the isomorphs of class 15 from the network:
```{r message=FALSE}
# This is the first one in the list
motifs_class_15 <- subgraph(wsnet, vids = iso_class_15_list[[1]]) |>
  as_sfnetwork()

# Join every other isomorph in the list
for(i in 2:length(iso_class_15_list)){
  motifs_class_15 <- st_network_join(motifs_class_15,
                   subgraph(wsnet, vids = iso_class_15_list[[i]]) |>
                     as_sfnetwork())
}
```

Plot the motifs of class 10:
```{r}
ggplot() + 
  geom_sf(data = walksheds_da |>
            filter(GeoUID == ID)) +
  geom_sf(data = wsnet |>
            activate(edges) |>
            st_as_sf()) +
  geom_sf(data = motifs_class_10 |>
            activate(edges) |>
            st_as_sf(),
          color = "red") +
  geom_sf(data = motifs_class_15 |>
            activate(edges) |>
            st_as_sf(),
          color = "blue") +
  theme_void()
```

Almost all elements of the network are part of a motif. But the number of motifs is low, which suggests that a large number of dead ends tends to reduce the number of motifs.

Degree distribution:
```{r}
ggplot(data = data.frame(degree = degree(wsnet)) |>
         mutate(degree = factor(degree,
                                ordered = TRUE))|>
         group_by(degree) |>
         summarize(count = n())) +
  geom_col(aes(x = degree, y = count))
```

Check walksheds with high number of motifs (normalized):
```{r}
d |> filter(normalized_motifs_3 > 1.07)
```

Choose a walkshed with high number of motifs:
```{r}
ID <- "35250675" #

wsnet <- hamilton_edges |>
  st_filter(
    walksheds_da |> filter(GeoUID == ID),
    .predicate = st_intersects)

# Convert to sfnetwork
wsnet <- as_sfnetwork(wsnet)
```

Find the motifs of class 10 and 15 in wsnet:
```{r}
iso_class_10_list <- subgraph_isomorphisms(size_3_iso[[10]], 
                                           wsnet, 
                                           method = "vf2")

iso_class_15_list <- subgraph_isomorphisms(size_3_iso[[15]], 
                                           wsnet, 
                                           method = "vf2")
```

Extract the isomorphs of class 10 from the network:
```{r message=FALSE}
# This is the first one in the list
motifs_class_10 <- subgraph(wsnet, vids = iso_class_10_list[[1]]) |>
  as_sfnetwork()

# Join every other isomorph in the list
for(i in 2:length(iso_class_10_list)){
  motifs_class_10 <- st_network_join(motifs_class_10,
                   subgraph(wsnet, vids = iso_class_10_list[[i]]) |>
                     as_sfnetwork())
}
```

Extract the isomorphs of class 15 from the network:
```{r message=FALSE}
# This is the first one in the list
motifs_class_15 <- subgraph(wsnet, vids = iso_class_15_list[[1]]) |>
  as_sfnetwork()

# Join every other isomorph in the list
for(i in 2:length(iso_class_15_list)){
  motifs_class_15 <- st_network_join(motifs_class_15,
                   subgraph(wsnet, vids = iso_class_15_list[[i]]) |>
                     as_sfnetwork())
}
```

Plot the motifs of class both classes:
```{r}
ggplot() + 
  geom_sf(data = walksheds_da |>
            filter(GeoUID == ID)) +
  geom_sf(data = wsnet |>
            activate(edges) |>
            st_as_sf()) +
  geom_sf(data = motifs_class_10 |>
            activate(edges) |>
            st_as_sf(),
          color = "red") +
  geom_sf(data = motifs_class_15 |>
            activate(edges) |>
            st_as_sf(),
          color = "blue") +
  theme_void()
```

Almost all elements of the network are part of a motif. But the number of motifs is low, which suggests that a large number of dead ends tends to reduce the number of motifs.

Degree distribution:
```{r}
ggplot(data = data.frame(degree = degree(wsnet)) |>
         mutate(degree = factor(degree,
                                ordered = TRUE))|>
         group_by(degree) |>
         summarize(count = n())) +
  geom_col(aes(x = degree, y = count))
```

```{r}
wsnet <- hamilton_edges |>
  st_filter(
    walksheds_da |> filter(GeoUID == ID),
    .predicate = st_intersects)

# Convert to sfnetwork
wsnet <- as_sfnetwork(wsnet)
```

Compute the degree distribution in the walksheds:
```{r warning=FALSE}
# Time algorithm
start.time <- Sys.time()

# Initialize a dataframe 
walksheds_da_degree_dist <- data.frame(GeoUID = numeric(), degree = numeric())

for(i in 1:nrow(walksheds_da)){
  wsnet <- hamilton_edges |>
    st_filter(
      walksheds_da[i,],
      .predicate = st_intersects)

  wsnet <- hamilton_edges |> 
    filter(edge_index %in% wsnet$edge_index)

  # Convert to sfnetwork
  wsnet <- as_sfnetwork(wsnet)
  walksheds_da_degree_dist <- rbind(walksheds_da_degree_dist,
                                    data.frame(GeoUID = walksheds_da$GeoUID[i],
                                               degree = degree(wsnet)))
}

Time difference:
```{r}
time.taken
```

Join urban fabric types:
```{r}
walksheds_da_degree_dist <- walksheds_da_degree_dist |> 
  left_join(walksheds_da |>
              st_drop_geometry(),
            by = "GeoUID")
```

Save object:
```{r}
save(walksheds_da_degree_dist, 
     file = "walksheds_da_degree_dist.rda", 
     compress = "xz")
```

```{r}
walksheds_da_degree_dist |>
  mutate(degree = factor(degree, 
                         ordered = TRUE)) |>
         group_by(Type, degree) |>
         summarize(count = n(),
                   .groups = "drop") |>
  ggplot() +
  geom_col(aes(x = degree, y = count, fill = Type), position = "dodge")
```

### 2023-07-06 data-raw/04-Accessibility-Calculations.Rmd
Updated analysis and files for paper.

@@

>>>>

Distribution of travel times by fabric type:
```{r}
ttm_walk_to_amenities |> 
  ggplot() +
  geom_histogram(aes(x = travel_time, fill = Type),
                 position = "dodge")
```

@@

See Millward, Spinney, and Scott (2013) for some distance-decay functions for active transport. The decay curves are very steep, so the difference in minimum time and mean time to destinations by urban fabric can potentially have a large impact on the decision to walk.

For all destinations this is the equation:
$$
\log(Y) = 1.282 - 0.048 \cdot t
$$
For all destinations this is the equation:
$$
\log(Y) = 1.035 - 0.032 \cdot t
$$

$Y$ is the percentage of trips and $t$ is the time in minutes.

```{r}
ggplot() +
  geom_function(fun = function(x) exp(1.282)*exp(-0.048)^x,
                color = "blue") +
  geom_function(fun = function(x) exp(1.035) * exp(-0.032)^x,
                color = "red") +
  #geom_vline(xintercept =  4.37) +
  #geom_vline(xintercept = 6.02) +
  xlim(1, 15)
```

The median of the minimum distance to any amenity in urban walksheds is 4.00 and in suburban sheds is 6.00. If we use the impedance function for all destinations, this means that the probability of taking an active trip is about 40\% lower in suburban sheds:
```{r}
exp(1.282)*exp(-0.048)^6.0/exp(1.282)*exp(-0.048)^4.0
```
